---
title: '[F] Always up to date: Using APIs as data source'
author: Sven
date: '2021-02-03'
slug: []
categories:
  - R
tags:
  - API
  - Data Analysis
---

# Intro

Hello again. Unfortunately exams are coming up very soon, which is why I was
so inactive lately. I was working on a project the entire time, but couldn't 
put as much effort in as I wanted to. But I will still post it anyway, because
it has been too long since my last post. 

# The project

On of the problems of today's data driven world is, that the data constantly 
gets updated. So how can you assure, that you're using the most up-to-date data?
That's where APIs come into play. I'm not going to explain what an API is and how
it works, because I want to keep this post as short as possible and focus on working 
with APIs in R, but if you're interested in learning more about APIs, you can follow this
[link](https://en.wikipedia.org/wiki/API){target="_blank"} to learn more about the topic.

I used [this](https://npgeo-corona-npgeo-de.hub.arcgis.com/datasets/dd4580c810204019a7b8eb3e0b329dd6_0){target="_blank"} as my data source, as Covid-19 is (unfortunately) a 
relevant topic at the moment. Before starting, we have to start as usual: By loading 
packages.

```{r}
#Loading packages
library(httr)
library(jsonlite)
```

These two packages are necessary to access web data in the JSON-format. 

```{r}
#API request
dt <- GET("https://services7.arcgis.com/mOBPykOjAyBO2ZKk/arcgis/rest/services/Coronaf%C3%A4lle_in_den_Bundesl%C3%A4ndern/FeatureServer/0/query?where=1%3D1&outFields=*&outSR=4326&f=json")
dt
```

In this step we're requesting the data from the API using the GET() function,
which is in the httr package. Then we can analyze the GET() output.
As you can see, the "Status:" shows the number 200, which is what we want to see,
as it corresponds to a successful attempt at getting the data. 

# THIS POST IS A WORK IN PROGRESS!!!








