---
title: '[F] Always up to date: Using APIs as data source'
author: Sven
date: '2021-02-15'
categories:
  - R
tags:
  - API
  - Data Analysis
---

# Intro

Hello again. Unfortunately exams are coming up very soon, which is why I was
so inactive lately. I was working on a project the entire time, but couldn't 
put as much effort in as I wanted to. But I will still post it anyway, because
it has been too long since my last post.

# The project

On of the problems of today's data driven world is, that the data constantly 
gets updated. So how can you assure, that you're using the most up-to-date data?
That's where APIs come into play. I'm not going to explain what an API is and how
it works, because I want to keep this post as short as possible and focus on working 
with APIs in R, but if you're interested in learning more about APIs, you can follow this
[link](https://en.wikipedia.org/wiki/API){target="_blank"} to learn more about the topic.

I used [this](https://npgeo-corona-npgeo-de.hub.arcgis.com/datasets/dd4580c810204019a7b8eb3e0b329dd6_0){target="_blank"} as my data source, as Covid-19 is (unfortunately) a 
relevant topic at the moment. Before starting, we have to start as usual: By loading 
packages.

```{r}
#Loading packages
library(httr)
library(jsonlite)
```

These two packages are necessary to access web data in the JSON-format. 

```{r}
#API request
dt <- GET("https://services7.arcgis.com/mOBPykOjAyBO2ZKk/arcgis/rest/services/Coronaf%C3%A4lle_in_den_Bundesl%C3%A4ndern/FeatureServer/0/query?where=1%3D1&outFields=*&outSR=4326&f=json")
dt
```

In this step we're requesting the data from the API using the GET() function,
which is in the httr package. Then we can analyze the GET() output.
As you can see, the "Status:" shows the number 200, which is what we want to see,
as it corresponds to a successful attempt at getting the data. 

In the next step we have to convert the data into the JSON-structure using the 
rawToChar() command from the jsonlite package.

```{r}
datCov <- fromJSON(rawToChar(dt$content))
names(datCov)
```

After converting the data to the JSON structure we can further explore our data.
This might take some trial and error, as I encountered some freezes when I 
tried to access the "wrong" data (i.e. not the actual data we're looking for).
So to reduce the risk of the site not deploying I will skip the detailed exploration
of the JSON data.

So the data we want is saved in "features" of our datCov object and again has multiple
types of data and information. To further extract the data we want we convert the 
"features" part to a list and finally convert our data into a data frame.

```{r}
listCov19 <- datCov$features
dfCov19 <- as.data.frame(listCov19[[1]])
```

Now we can also take our first look at the actual data we want to work with.

```{r}
dim(dfCov19)
head(dfCov19)
sapply(dfCov19, class)
```

The data frame has 18 columns and 16 rows and contains various data about 
Covid-19 in Germany. 

From here you can do whatever you want with the data. Use it for analysis or create
plots like those two!

```{r, echo=FALSE}
dfc <- dfCov19
dfc$LAN_ew_GEN <- c("SH", "HH", "NI", "HB", "NW", "HE", "RP", "BW",
                    "BY", "SL","BE", "BB", "MV", "SN", "ST", "TH")

library(ggplot2)
library(cowplot)

cp100k <- ggplot(dfc, aes(reorder(LAN_ew_GEN, -cases7_bl_per_100k), 
                     cases7_bl_per_100k)) + geom_bar(stat="identity") + 
  ggtitle("Covid-19 in Germany") + xlab("State (short)") +
  ylab("new cases per 100.000 inhabitants (7 days)")

human_numbers <- function(x = NULL, smbl ="", signif = 1){
  humanity <- function(y){
    
    if (!is.na(y)){
      tn <- round(abs(y) / 1e12, signif)
      b <- round(abs(y) / 1e9, signif)
      m <- round(abs(y) / 1e6, signif)
      k <- round(abs(y) / 1e3, signif)
      
      if ( y >= 0 ){
        y_is_positive <- ""
      } else {
        y_is_positive <- "-"
      }
      
      if ( k < 1 ) {
        paste0( y_is_positive, smbl, round(abs(y), signif ))
      } else if ( m < 1){
        paste0 (y_is_positive, smbl,  k , "k")
      } else if (b < 1){
        paste0 (y_is_positive, smbl, m ,"m")
      }else if(tn < 1){
        paste0 (y_is_positive, smbl, b ,"bn")
      } else {
        paste0 (y_is_positive, smbl,  comma(tn), "tn")
      }
    } else if (is.na(y) | is.null(y)){
      "-"
    }
  }
  
  sapply(x,humanity)
}

ct <- ggplot(dfc, aes(reorder(LAN_ew_GEN, -Fallzahl), 
                     Fallzahl)) + geom_bar(stat="identity") + 
  ggtitle("") + xlab("State (short)") +
  ylab("Total cases") +
  scale_y_continuous(labels = human_numbers)

plot_grid(cp100k, ct, labels="")
```

That's all there is for this blog post. Sorry again for not posting as much lately.
University is more challenging at the moment than I had anticipated. 
But soon all of my exams will be over and I promise I will post more!
As always you can find the code for the project on my [GitHub](https://github.com/SvenNekula/Covid_API){target="_blank"}.
Thank you for taking the time to read this post! 

See you next time! -SN 


