---
title: '[F] Crawling the web with Python'
author: 'Sven '
date: '2021-04-08'
slug: []
categories:
  - Python
tags:
  - webcrawler
---

Hello again! Unfortunately I caught the corona virus a few weeks back and thus 
had to stay in quarantine for ~2 weeks. Quarantine sucks, but at least I used 
the time to get my python skills back up to par. 

# The project

I wanted to create a webcrawler using python. Being a fun project, I used a website with data from all the pokemon to crawl the data and write it to a csv file. 
During the process I found two ways to do so.
The first way required more code than the second one. But I will show you both.

# V1 

As usual we have to load the modules first.

```{python, eval=FALSE}
import requests
from bs4 import BeautifulSoup
```

Next up we have to set the url and use the request package to make a http request.

```{python, eval=FALSE}
url = 'https://pokemondb.net/pokedex/all'
r = requests.get(url)
```

You can check if the request was successfull by printing the status code using
the command print(r.status_code). If the status-code is 200 you're good. After 
creating the request and checking if it was successfull we have to use the 
'BeautifulSoup' module to parse the website's html file. 

```{python, eval=FALSE}
soup = BeautifulSoup(r.text, 'html.parser')
```

After parsing the html file we can start to look for the data. In our case the 
data is in a html-table, which we can find using the .find command.

```{python, eval=FALSE}
pokemon_table = soup.find('table', class_ = 'data-table block-wide')
```

Now that we found the html-table we can make a list using a for loop. The 
columns of the table have different classes however, which we can find by 
inspecting the website using our web-browser. There were 3 different classes:
'cell-name', 'cell-total' and 'cell-num'. There were 7 different elements of the 
'cell-num'-class, so we had to further inspect those elements, to get the right ones.
Having done that, we can create the list. 

```{python, eval=FALSE}
pokemon_list = []
for pkmn in pokemon_table.find_all('tbody'):
    rows = pkmn.find_all('tr')
    for row in rows:
        p_name = row.find('td', class_ = 'cell-name').text
        p_total = row.find('td', class_ = 'cell-total').text
        p_hp = row.find_all('td', class_ = 'cell-num')[1].text
        p_att = row.find_all('td', class_ = 'cell-num')[2].text
        p_def = row.find_all('td', class_ = 'cell-num')[3].text
        p_spatk = row.find_all('td', class_ = 'cell-num')[4].text
        p_spdef = row.find_all('td', class_ = 'cell-num')[5].text
        p_speed = row.find_all('td', class_ = 'cell-num')[6].text
        pokemon_list.append([p_name, p_total, p_hp, p_att, p_def, p_spatk, p_spdef, p_speed])
```

The last step is to convert this list into a csv-file. We can do this with the help
of the csv module.

```{python, eval=FALSE}
import csv
with open('allpokemon.csv', 'w', newline='') as csvfile:
    pkmnwriter = csv.writer(csvfile, delimiter=';',
                            quotechar='""', quoting=csv.QUOTE_MINIMAL)
    
    for row in pokemon_list:
        pkmnwriter.writerow
```

Now we successfully transformed the website's data into a csv-file!

# V2

The second way of transforming the website's data into a csv is a lot easier.
It only requires 4 lines of code, because we're using pandas' read_html function.
This function automatically scans the website for html-tables and converts them to 
a list. One thing to keep in mind is to use the requests module the same way we used 
it in V1, otherwise you might encounter some errors!

```{python, eval=FALSE}
import pandas as pd

pkmn_list = pd.read_html(r.text)
df = pd.DataFrame(pkmn_list[0])

df.to_csv (r'P:\\GitHub\\webcrawler\\pokemon_list.csv', index = False, header=True)
```

Those were 2 ways to crawl data from a website. Now we can further use the data
and analyze it. 

# Analyzing the data

Before analyzing the data I opened the csv file in excel and transformed it to 
a .xslx file, because it makes it easier to work with (in excel).
To load the .xslx file into RStudio I used the 'readxl' package. I also used 
'dplyr' and 'ggplot2' to filter and visualize the data. 

```{r, echo=FALSE, warning=FALSE}
#loading packages
library(readxl)
library(dplyr)
library(ggplot2)

#loading data
pData <- read_excel("P:\\GitHub\\webcrawler\\pokemon_list.xlsx")
```

Now we can get an overview of the data.

```{r}
#overview of data
str(pData)
```

As you can see the data is the 'tibble'-format. I wanted to change that and turn 
the data into a dataframe. To do that I used the as.data.frame function.

```{r}
#converting tibble into df
dfp <- as.data.frame(pData)
str(dfp)
```

If you have knowledge of Pokemon, you probably know that Eternamax is the strongest
Pokemon. However it's just used in an NPC fight and can't be used by the player,
which is why it has to be removed from the dataset, because it's an extreme outlier.
This can be done multiple ways, I just chose to use the subset function, as I've 
never used it. 

```{r}
#removing Eternamax
max(dfp$Total)
pokemon_data <- subset(dfp, dfp$Total<1125)
```

After finalizing the dataframe I wanted to work with, I wrote a function called
'quickAnalysis'. Pokemon have different stats, for example attack or defense.
This function returns a list with the mean, median, max and min of a stat. 
The function takes 2 arguments: x, i.e. the column of the dataframe and col, i.e.
the perticular column. I chose to do it this way, so that the function can also return
the name of the stat.

```{r}
quickAnalysis <- function(x, col){
  v_mean <- mean(x)
  v_median <- median(x)
  v_max <- max(x)
  v_min <- min(x)
  ql <- list(v_mean, v_median, v_max, v_min)
  names(ql) <- c(paste("Mean of", names(dfp)[col], "Stat"), 
                 paste("Median of", names(dfp)[col], "Stat"), 
                 paste("Max of", names(dfp)[col], "Stat"),
                 paste("Min of", names(dfp)[col], "Stat"))
  return(ql)
}
```

We can plug in the different stats into the function now. 

```{r}
#total
quickAnalysis(pokemon_data[,4], 4)
#hp
quickAnalysis(pokemon_data[,5], 5)
#att
quickAnalysis(pokemon_data[,6], 6)
#def
quickAnalysis(pokemon_data[,7], 7)
#spatk
quickAnalysis(pokemon_data[,8], 8)
#spdef
quickAnalysis(pokemon_data[,9], 9)
#speed
quickAnalysis(pokemon_data[,10], 10)
```

As you can see the max total stat of any Pokemon is 780, while the min is 175.
That's quite a big margin. To visualize the distribution of the Pokemon totals we can
use a histogram. 

```{r}
#plotting data:
ht <- ggplot(pokemon_data, aes(x=Total)) + geom_histogram(binwidth = 10, color="black", fill="white") +
  scale_x_continuous(breaks = seq(0, 1150, 50)) +
  geom_vline(xintercept = mean(pokemon_data$Total), col="blue", linetype = "dashed", size = 1) +
  geom_density(fill="red", alpha=.2, aes(y=10 * ..count..,))

ht
```

Here we can see the mean of the total stat, which is ~440. We can also see that 
the most common total base stats are between 480 and 500. 


But what are the best and the worst Pokemon? To find this out we can combine 
our quickAnalysis function with dplyr's filter function.

```{r}
best_pkmn <- pokemon_data %>%
  filter(Total==unlist(quickAnalysis(pokemon_data[,4], 4)[3]))
best_pkmn[,2:4]
```

The best Pokemon are the mega evolutions of Mewtwo and Rayquaza with a base stat total of 780.

```{r}
worst_pkmn <- pokemon_data %>%
  filter(Total==unlist(quickAnalysis(pokemon_data[,4], 4)[4]))
worst_pkmn[,2:4]
```

And the worst Pokemon is Wishiwashi's solo form with a base stat total of 175.


I also wanted to make a boxplot showing the different stats. To do that we had 
to reshape our dataframe from a wide-format to a long-format. 

```{r}
#reshaping data to plot
plotDF <- data.frame(type=c(rep(names(pokemon_data)[5], 1044), 
                            rep(names(pokemon_data)[6], 1044),
                            rep(names(pokemon_data)[7], 1044),
                            rep(names(pokemon_data)[8], 1044),
                            rep(names(pokemon_data)[9], 1044),
                            rep(names(pokemon_data)[10], 1044)),
                     value=c(pokemon_data[,5],
                             pokemon_data[,6],
                             pokemon_data[,7],
                             pokemon_data[,8],
                             pokemon_data[,9],
                             pokemon_data[,10]))

head(plotDF)
```

Now we are able to create a boxplot.

```{r}
d_stats <- ggplot(plotDF, aes(x=value, y=type, color=type)) +
  geom_boxplot()
d_stats
```

We can see again that the attack stat has the highest median. 
It's also visible that the HP stat has a lot of outliers. 


As always you can find the code for the [crawler](https://github.com/SvenNekula/webcrawler){target="_blank"} and the 
[analysis](https://github.com/SvenNekula/CrawlerAnalysis){target="_blank"} on 
my GitHub.
That's it for this post. Thanks for reading, see you again next time!

-SN




